---
title: "IWT CRC"
author: "Davit Sargsyan"
date: "11/24/2021"
output:
  html_document:
    df_print: paged
---

```{r setup}
# Install qvalue package from bioconductor
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("qvalue")

require(data.table)
require(ggplot2)
require(parallel)
require(glmnet)
```

# Load data
```{r}
setwd("/cloud/project")
load("Data/otu_431_clr.Rda")
load("Data/rna_crcad.Rda")
load("Data/group_crcad.Rda")
```

# RNA
## Merge disease status with RNA expression
```{r}
rna <- data.table(id = rownames(rna.crcad),
                  rna.crcad)

tmp <- data.table(id = names(group),
                  disease = group)

rna <- merge(tmp,
             rna,
             by = "id",
             all = TRUE)
rna$disease <- factor(rna$disease)
summary(rna$disease)
```

```{r}
ggplot(rna,
       aes(x = disease,
           y = `100132116_at`,
           group = id)) +
  geom_point(position = position_dodge(0.3)) +
  theme_bw() +
  theme(legend.position = "none")
```

## Step 1: univariable p-values
```{r}
univar <- lapply(rna[, 3:ncol(rna)],
      FUN = function(a) {
        # t.test(a ~ rna$disease,
        #        alternative = "two.sided",
        #        var.equal = FALSE)$p.value
        s1 <- summary(glm(rna$disease ~ a,
            family = "binomial"))
        return(s1$coefficients[2, 4])  #obtained the p-values
      })
univar <- do.call("c", univar)


# Problem with the margins (too large)
par(mar=c(1,1,1,1))
hist(univar, 20)
summary(univar)
```

## Step 2: weights
```{r}
q <- qvalue::qvalue(univar)$qvalues
wgt <- -log(q)
wgt <- wgt/sum(wgt)

# Problem with the margins (too large)
par(mar=c(1,1,1,1))
hist(wgt, 20)
```

## Step 3: subset of features
### Create a cluster of cores
```{r}
ncores <- detectCores()
ncores
cl <- makeCluster(getOption("cl.cores",
                            4))

# NOTE: export data AND packages to cluster!
# clusterExport(cl = cl, 
#               varlist = c("rna",
#                           "wgt",
#                           "alpha"))
clusterExport(cl = cl, 
              varlist = c("rna",
                          "wgt"))

cl
```

### Main algorithm
**NOTES:**  
1. I did not split the data into training/testing sets. Previously, I used 2/3 of subjects to build the model and the other 1/3 to check model fit.  
   
```{r, cache=TRUE}
system.time({
  out <- parLapply(cl = cl,
                   X = 1:1000,
                   fun = function(i) {
                     require(data.table)
                     require(glmnet)
                     
                     # sample 2/3 of the subjects at random
                     id_keep <- sample(x = rna$id,
                                       size = floor(2*nrow(rna)/3),
                                       replace = FALSE)
                     
                     # sample sqrt(features), weighted
                     feat_keep <- sample(x = 3:ncol(rna),
                                         size = floor(sqrt(ncol(rna) - 2)),
                                         replace = FALSE,
                                         prob = wgt)
                     
                     
                     # subset data
                     dt1 <- rna[id %in% id_keep,
                                c(2,
                                  feat_keep),
                                with = FALSE]
                     
                     # glmnet
                     try({
                       # get lambda min
                       m0 <- cv.glmnet(x = as.matrix(dt1[, -1]),
                                       y = dt1$disease,
                                       family = "binomial",
                                       alpha = 0.5,
                                       nfolds = 5)
                       
                       # use lambda min in glmnet
                       m1 <- glmnet(x = as.matrix(dt1[, -1]),
                                    y = dt1$disease,
                                    family = "binomial",
                                    alpha = 0.5,
                                    lambda = m0$lambda.min)
                       
                       feat_selected <- rownames(m1$beta)[as.numeric(m1$beta) != 0]
                     })
                     
                     if (length(feat_selected) > 0) {
                       
                       
                       return(list(status = "OK",
                                   feat_sampled = rownames(m1$beta),
                                   feat_sampled_all = data.table(feat = colnames(rna)[3:ncol(rna)],
                                                                 feat_sampled = colnames(rna)[3:ncol(rna)] %in%
                                                                   rownames(m1$beta)),
                                   feat_selected = feat_selected,
                                   feat_selected_all = data.table(feat = colnames(rna)[3:ncol(rna)],
                                                                  feat_selected = colnames(rna)[3:ncol(rna)] %in% feat_selected),
                                   dev_ratio = m1$dev.ratio))
                     } else {
                       return(status = NA)
                     }
                   })
})
```

## stop the cluster
```{r}
stopCluster(cl)
gc()
```
## Remove empty iterations
```{r}
tmp <- lapply(out,
              function(a) {
                is.na(a$status)
              })
tmp <- do.call("c", tmp)
out[tmp == TRUE] <- NULL
```

## Number of times each feature was sampled
### Unweighted
```{r, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}
tmp <- lapply(out,
              function(a){
                as.data.frame(a$feat_sampled_all) 
              })


dt_feat_sampled <- Reduce(function(x, y)  {merge(x, y, all = TRUE, by = "feat")}, tmp)
```

### Weighted
```{r, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}
tmp <- lapply(out,
              function(a){
                df1 <- as.data.frame(a$feat_sampled_all) 
                df1$feat_sampled_wgt <- df1$feat_sampled*a$dev_ratio
                df1$feat_sampled <- NULL
                return(df1)
              })


dt_feat_sampled_wgt <- Reduce(function(x, y)  {merge(x, y, all = TRUE, by = "feat")}, tmp)
```

### Merge
```{r, cache=TRUE}
t1 <- data.table(id = dt_feat_sampled[, 1],
                 feat_sampled = rowSums(dt_feat_sampled[, -1]))

t2 <- data.table(id = dt_feat_sampled_wgt[, 1],
                 feat_sampled_wgt = rowSums(dt_feat_sampled_wgt[, -1]))

t12 <- merge(t1, t2, by = "id")
```


## Number of times each feature was selected
### Unweighted
```{r, warning=FALSE,message=FALSE,echo=FALSE}
tmp <- lapply(out,
              function(a){
                as.data.frame(a$feat_selected_all) 
              })


dt_feat_selected <- Reduce(function(x, y)  {merge(x, y, all = TRUE, by = "feat")}, tmp)
```

### Weighted
```{r, warning=FALSE, message=FALSE, echo=FALSE, cache=TRUE}
tmp <- lapply(out,
              function(a){
                df1 <- as.data.frame(a$feat_selected_all) 
                df1$feat_selected_wgt <- df1$feat_selected*a$dev_ratio
                df1$feat_selected <- NULL
                return(df1)
              })


dt_feat_selected_wgt <- Reduce(function(x, y)  {merge(x, y, all = TRUE, by = "feat")}, tmp)
```

### Merge
```{r, cache=TRUE}
t3 <- data.table(id = dt_feat_selected[, 1],
                 feat_selected = rowSums(dt_feat_selected[, -1]))

t4 <- data.table(id = dt_feat_selected_wgt[, 1],
                 feat_selected_wgt = rowSums(dt_feat_selected_wgt[, -1]))



t34 <- merge(t3, t4, by = "id")
```


# Feature importance
```{r, cache=TRUE}
res <- data.table(merge(t12, t34, by = "id"))


res[, imp := feat_selected/feat_sampled]
res[, imp_wgt := feat_selected_wgt/feat_sampled_wgt]


setorder(res, -imp)
head(res[imp != "NaN",])

# write.csv(res,
#           file = "Output/res.csv",
#           row.names = FALSE)
```


# Get the selected gene names
```{r, cache=TRUE}
# Install biobase
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("Biobase")

# Load data
load("Data/esetRna_new.Rda")


fdata <- as(featureData(esetRna), "data.frame")
temp <- subset(fdata, select = c(SYMBOL))
res <- base::merge(as.data.frame(res), temp, by.x="id", 
                   by.y="row.names", all.x=TRUE)
names(res)[names(res) == "SYMBOL"] <- "name"

res <- res[order(-res$imp_wgt), ]

# Save the res data
write.csv(res, file="Output/res_uArray.csv",
          row.names=FALSE)

```






# Session
```{r session}
sessionInfo()
```

